---
title: "DSC Assignment 6"
output:
  word_document: default
  pdf_document: default
  html_document: default
---
```{r}
setwd("~/Desktop/class/dsc")
asthma<-read.csv("asthma.csv")
```

# 1-a why wilcoxon?
```{r}
hist(asthma$air)
hist(asthma$sulf.diox)
hist(asthma$air - asthma$sulf.diox)
```
Wilcoxon Signed-Rank Test does not require populations to be normally distributed as the paranetric paried t-test did. Based on the histograms, they do not appear to be normally distributed, so we want to use Wilcoxon signed rank test

#1-b by hands
```{r}
a<-cbind(asthma$air-asthma$sulf.diox)
sort(a)
sum_positive<-2+3+6+10; sum_positive
sum_negative<-1+4+5+7+8+9+11+12+13+14+15+16+17; sum_negative
# T=21, n=17
Mt<-17*18/4;Mt
sigmat<-sqrt(17*18*35/24);sigmat
Zt<-(21-76.5)/sigmat;Zt
2*pnorm(-2.672)
```
p-value is 0.0075, under a=0.05 the null hypothesis is rejected

#1-c 
```{r}
wilcox.test(asthma$air, asthma$sulf.diox, paired=T, exact=F, correct=F)

```

#1-d
```{r}
wilcox.test(asthma$air, asthma$sulf.diox, paired=T, exact=T, correct=F)

0.008707-0.006653
```
0.002054 differ from result in part c 

#2-a
```{r}
bulimia<-read.csv("bulimia.csv")
bulimic<-bulimia$bulimic
healthy<-bulimia$health[1:15]
```

#2-b
```{r}
hist(bulimic)
hist(healthy)
```
Based on the histograms, the two samples do not appear to be coming from normally distributed populations, so we want to use the Wilcoxon rank sum test

#2-c
```{r}
sort(bulimic)
sort(healthy)
rank_bul<-1+2+3+4+5+6+7+8+9+10+12+13+14+16+18+20+21.5+23+25+27+28+29+32; rank_bul
rank_heal<-11+15+17+19+21.5+24+26+30.5+30.5+33+34+35+36+37+38; rank_heal
W<-333.5; n1<-23;n2<-15
Mw<-n1*(n1+n2+1)/2;Mw
sigmaw<-sqrt(n1*n2*(n1+n2+1)/12);sigmaw
Zw<-(W-Mw)/sigmaw;Zw; pnorm(-3.4343)*2;
```
p-value is 0.005940, under a=0.05, the null hypothesis is rejected.
#2-d
```{r}
wilcox.test(bulimia$bulimic, bulimia$healthy, exact=F,correct=F)
```

# 3-a
```{r}
actg<-read.csv("actg.csv")
actg$IVDrug<-as.factor(actg$IVDrug)
boxplot(actg$Age~actg$IVDrug)
```
no

#3-b
```{r}
library(dplyr)
cu<-actg %>%
  filter(actg$IVDrug =="Currently")
ne<-actg %>%
  filter(actg$IVDrug =="Never")
pre<-actg %>%
  filter(actg$IVDrug =="Previously")
n1<-28; n2<-944;  n3<-179
mean(actg$Age)
var_cu<-sd(cu$Age)^2
var_ne<-sd(ne$Age)^2
var_pre<-sd(pre$Age)^2
s2w<-(27*var_cu +943*var_ne+178*var_pre)/(28+944+179-3)
s2b<-(n1*((mean(cu$Age)-38.61251)^2)+n2*((mean(ne$Age)-38.61251)^2)+n3*((mean(pre$Age)-38.61251)^2))/2
F<-s2b/s2w;F
1-pf(72.451,2,1148)
```
under a=0.05, pvalue=0,the null hypothesis is rejected.


#3-c
```{r}
model1<-aov(actg$Age~actg$IVDrug)
anova(model1)
```
# d)
```{r}
pairwise.t.test(actg$Age, actg$IVDrug,p.adj="bonferroni")
```


















