---
title: "DSC Homework 7"
output:
  pdf_document: default
  html_document: default
---

```{r}
setwd("~/Desktop/class/dsc")
house<-read.csv("houses.csv", header=T)
```

#1 state assumptions in terms of the variables
# y= sellingprice ~ x= livingsize
```{r}
# 1 : given living sizes, the selling prices are independent.
# 2 : There is a linear relationship between selling prices and living sizes( Expected value of error terms = 0)
# 3 : The variance is constant across all values of living sizes, known as homoscedasticity.
# 4 : For a specified value of selling prices and living sizes is normally distributed.
# 5 : living sizes are fixed, known quantities.
```

# 2 pearson 
```{r}
cor(house$livingsize, house$sellingprice)
```
0.626

#3 scatter plot of selling price over living space size
```{r}
plot(house$livingsize, house$sellingprice)
```
It seems these two variables have a positive relationship. So linear regression would be good for it.

#4 simple regression, report the prediction equation
```{r}
house.lm<- lm(sellingprice~livingsize, data=house)
house.lm
```

selling price = 15.30 + 14.01 * (living size)

#5 estimated mean selling price with living space size 1.5
```{r}
predict(house.lm, newdata=data.frame(livingsize=1.5))

```
36.30467

# 6 space size of 3.2
```{r}
predict(house.lm, newdata=data.frame(livingsize=3.2))
```
60.1154

# 7 Yes. Those two questions have the same equation and what I need to do is plugging in the different x-values into the same formula.

# 8 1.2 feet 90% CI
```{r}
predict(house.lm, newdata=data.frame(livingsize=1.2), interval="confidence", level=0.9)
```
[30.15566, 34.0499]
the probability that selling price for living size=1.2 is between 30.15566 and 34.0499 is 90 %

# 9 90% predicition, selling price for 1.2
```{r}
predict(house.lm, newdata=data.frame(livingsize=1.2), interval="predict", level=0.9)
```
90 % confidence that the interval (23.69295, 40.51261) contains the true selling prices when the living size is at 1.2

# 10 construct 95% confidence interval for the slope of the estimated regression equation and interpret the results.
```{r} 
# n=26 B1=14.01
s2<-1/24*sum((house$sellingprice-mean(house$sellingprice))^2)
denominator<-sum((house$livingsize-mean(house$livingsize))^2)
varB1<-s2/denominator; SEB1<-sqrt(varB1)

qt(0.975,24)
(14.01-2.063899*SEB1);(14.01+2.063899*SEB1)
```
[4.583954, 23.43605]
95% confident that that interval between 4.583954 and 23.43605 captures the true population slope

# 11 Test B1=0 or not a=0.05
```{r}
14.01/SEB1 # t
2*(1-pt(3.067, df=14)) #p-value
```
Since the p-value is less than 0.05, we reject the null hypothesis and conclude that there is a significant linear relationship between selling price and living size.

#12 what is the value of the coefficient determination
```{r}
summary(house.lm)
```
Coefficient of Determination is 0.3919
R^2 represents the proportion of variability in selling price that is explained by its linear relationship with livingsize. Since our R^2 is 0.3919, there is positive linear relationship.

#  13 diagnostic plot (residual plot and normal qq plot), comment vvv
```{r}
par(mfrow=c(2,2))
plot(house.lm)
```


#14 calculate the prediction equation for selling price as a function of living space size and whether or not the house has a fireplace. report the prediction equatiuon. at the a=0.10, does having fireplace significantly predict the selling price once we have accounted for living space size?
```{r}
house.lm1<-lm(sellingprice~livingsize+fireplace, data=house)
summary(house.lm1)
```
Selling Price = 14.630 + 13.711*(livingsize)+ 3.987*(fireplaceyes)
 
P-value is 0.057660, therefore under a=0.1, we reject the null hypothesis that B1(firespace)=0. Therefore having fireplace increases predicting selling price.( R-squares now 0.4818 from 0.3919)

R-square 0.4818


#15
```{r}
house.lm2<-lm(sellingprice~livingsize*fireplace, data=house)
summary(house.lm2)
```
selling price= 13.993+ 14.174*(livingsize)+ 6.073*(fireplace) -1.498(living & fireplace)

the interaction is not that significant at a=0.05.
Because we can see the t-value of interaction is -0.202. which is small.

R-sqaure 0.4828
#16

```{r}
house.lm3<-lm(sellingprice~livingsize+fireplace+age+sitesize, data=house)
summary(house.lm3)
```
sellingprice = 22.517 + 12.97*livingsize +5.07* fireplace -0.17466 * Age -0.11566 * Sitesize

R-square 0.6219

#17
```{R}
aov(house.lm3)
house.lm4<-lm(sellingprice~livingsize+fireplace, data=house)
aov(house.lm4)


F<-((467.6292-353.6574)/(4-2))/ (341.1998/(26-4-1))
1-pf(3.507341, 21,23)
```
p-value is 0.0021 therefore adding the two variables in model increases model significantly. 
F-statistic= 3.50












